### 2025-10-13 poster/old/neon_poster_airi.tex
finished poster

% Template file for an a0 landscape poster.
% Written by Graeme, 2001-03 based on Norman's original microlensing
% poster.
%
% See discussion and documentation at
% <http://www.astro.gla.ac.uk/users/norman/docs/posters/> 
%
% $Id: poster-template-landscape.tex,v 1.2 2002/12/03 11:25:46 norman Exp $

% Default mode is landscape, which is what we want, however dvips and
% a0poster do not quite do the right thing, so we end up with text in
% landscape style (wide and short) down a portrait page (narrow and
% long). Printing this onto the a0 printer chops the right hand edge.
% However, 'psnup' can save the day, reorienting the text so that the
% poster prints lengthways down an a0 portrait bounding box.
%
% 'psnup -w85cm -h119cm -f poster_from_dvips.ps poster_in_landscape.ps'

\documentclass[a0]{a0poster}
% You might find the 'draft' option to a0 poster useful if you have
% lots of graphics, because they can take some time to process and
% display. (\documentclass[a0,draft]{a0poster})
\input defs
\pagestyle{empty}
\setcounter{secnumdepth}{0}
\renewcommand{\familydefault}{\sfdefault}
\newcommand{\QED}{~~\rule[-1pt]{8pt}{8pt}}\def\qed{\QED}

\renewcommand{\reals}{{\mbox{\bf R}}}

% The textpos package is necessary to position textblocks at arbitary 
% places on the page.
\usepackage[absolute]{textpos}

\usepackage{fleqn,psfrag,wrapfig,tikz}
\usepackage{enumitem}
\usepackage{amsmath}

\usepackage{caption}

%\usepackage{svg}
\usepackage[papersize={38in,28in}]{geometry}

% Graphics to include graphics. Times is nice on posters, but you
% might want to switch it off and go for CMR fonts.
\usepackage{graphics}

% we are running pdflatex, so convert .eps files to .pdf
%\usepackage[pdftex]{graphicx}
%\usepackage{epstopdf}

% Russian language support
\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}

% These colours are tried and tested for titles and headers. Don't
% over use color!
\usepackage{color}
\definecolor{Red}{rgb}{0.9,0.0,0.1}

\definecolor{bluegray}{rgb}{0.15,0.20,0.40}
\definecolor{bluegraylight}{rgb}{0.35,0.40,0.60}
\definecolor{gray}{rgb}{0.3,0.3,0.3}
\definecolor{lightgray}{rgb}{0.7,0.7,0.7}
\definecolor{darkblue}{rgb}{0.2,0.2,1.0}
\definecolor{darkgreen}{rgb}{0.0,0.5,0.3}

\renewcommand{\labelitemi}{\textcolor{bluegray}\textbullet}
\renewcommand{\labelitemii}{\textcolor{bluegray}{--}}

\setlength{\labelsep}{0.5em}

% see documentation for a0poster class for the size options here
\let\Textsize\normalsize
%\def\Head#1{\noindent\hbox to \hsize{\hfil{\LARGE\color{bluegray} #1}}\bigskip}
\def\Head#1{\noindent{\LARGE\color{black} #1}\bigskip}
\def\LHead#1{\noindent{\LARGE\color{bluegray} #1}\bigskip}
\def\Subhead#1{\noindent{\large\color{bluegray} #1}\bigskip}
\def\Title#1{\noindent\hbox to \hsize{\hfil{\fontsize{72}{86}\selectfont\color{Red} #1}}}

% Set up the grid
%
% Note that [40mm,40mm] is the margin round the edge of the page --
% it is _not_ the grid size. That is always defined as 
% PAGE_WIDTH/HGRID and PAGE_HEIGHT/VGRID. In this case we use
% 23 x 12. This gives us three columns of width 7 boxes, with a gap of
% width 1 in between them. 12 vertical boxes is a good number to work
% with.
%
% Note however that texblocks can be positioned fractionally as well,
% so really any convenient grid size can be used.
%
\TPGrid[40mm,40mm]{23}{12}      % 3 cols of width 7, plus 2 gaps width 1

\parindent=0pt
\parskip=0.2\baselineskip

\begin{document}

% Understanding textblocks is the key to being able to do a poster in
% LaTeX. In
%
%    \begin{textblock}{wid}(x,y)
%    ...
%    \end{textblock}
%
% the first argument gives the block width in units of the grid
% cells specified above in \TPGrid; the second gives the (x,y)
% position on the grid, with the y axis pointing down.

% You will have to do a lot of previewing to get everything in the 
% right place.

% AIRIxTGU logo in the top left corner
\begin{textblock}{2}(0,0)
\includegraphics[width=4\TPHorizModule]{AIRIxTGU_logos.pdf}
\end{textblock}

% "Лето с AIRI" in the top right corner
\begin{textblock}{5}(19.5, 0)
    \begin{center}
    {\Large \textcolor{bluegray}{\textbf{Лето с AIRI}}}
    \end{center}
    \end{textblock}

% This gives good title positioning for a portrait poster.
% Watch out for hyphenation in titles - LaTeX will do it
% but it looks awful.
\begin{textblock}{23}(0,-0.1)
    \begin{center}
    {\fontsize{96}{110}\selectfont\color{Red} \textbf{Neon: Nuclear Norm to Beat Muon}}
    \end{center}
    \end{textblock}

\begin{textblock}{23}(0,0.2)
    \begin{center}
    {\large Alexey Kravatskiy, Ivan Kozyrev, Nikolay Kozlov, Alexander Vinogradov, Daniil Merkulov, Ivan Oseledets}

    \vspace{0.5em}

    {\normalsize \color{bluegray} \emph{Optimization Class Project, MIPT}}
    \end{center}
    \end{textblock}

% Vertical lines to separate sections - very thin and grey
\begin{textblock}{0.02}(7.5,1.0)
\textcolor{lightgray}{\rule{0.02\TPHorizModule}{10\TPVertModule}}
\end{textblock}

\begin{textblock}{0.02}(15.5,1.0)
\textcolor{lightgray}{\rule{0.02\TPHorizModule}{10\TPVertModule}}
\end{textblock}

\begin{textblock}{7.0}(0,1.25)

\Head{Introduction}\\
Recent advances in optimization techniques have highlighted the benefits of leveraging the matrix structure of neural network weights during training. Optimizers like Muon (Jordan et al.) and Shampoo (Gupta, Koren and Singer) have shown promise, but at a significantly higher computational cost than traditional methods like Adam. To bridge this gap, we propose Neon, a new optimizer that builds upon the framework of Bernstein and Newhouse. By using alternative norms, such as nuclear norm (Nuclear-Neon) or custom $F*$ norm ($F*$-Neon), we obtain low-rank update matrices that enable more efficient computation. We evaluate the performance of Neon, Muon, and Adam on training MLP and CNN on CIFAR-10 and on finetuning NanoGPT.

\medskip
\Head{Neon's update rule}\\
Bernstein and Newhouse suggest obtaining the update step for a weight matrix $W$ as a solution to the optimization problem:
\begin{equation}
    \langle G, \delta W \rangle + \frac{\lambda}{2} \norm{\delta W}^2 \to \min_{\delta W}\,,
\end{equation}
where $G$ is a gradient-like matrix obtained via backpropagation. Setting norm to RMS-to-RMS norm (scaled version of Spectral norm) produces Muon. We consider two different choices instead:
\begin{enumerate}
    \item Choosing nuclear norm, $\norm{\cdot}_*$, produces rank-1 update, defined by 
    \begin{equation}\label{eqn:update_star}
        \delta W = -\frac{1}{2\lambda} u_1 \sigma_1 v_1^T\,,
    \end{equation}
    where $\sigma_1$ is largest singular value of $G$, and $u_1$, $v_1$ are corresponding singular values.

    \item Choosing $F*$ norm, defined by $\norm{\cdot}_{F*} = (\norm{\cdot}_{*} + \norm{\cdot}_{F})/2$, produces a relatively small-rank update, defined by
    \begin{equation}\label{eqn:update_F_star}
    \delta W = -\frac{1}{\lambda}UDV^T
    \end{equation} 
    with $D = \diag(d_i)$, where $d_i = [\sigma_i - \tau]_+$ and $\tau$ is given by
    \begin{equation*}
        \sum_{i=1}^n [\sigma_i - \tau]_+ = \tau\,.
    \end{equation*}
\end{enumerate}

\medskip
\Head{Efficient update computation}\\
% A couple of sentences about Lanczos algorithm and reference computation times and required number of iterations for tipical value of $\tau$.
We use cupy's svds routine to obtain gradients' matrices' SVD approximation formed by largest singular values and corresponding vectors.
By applying the Lanczos process to either $A^TA$ or $AA^T$, it generates a sequence of orthogonal vectors (Lanczos vectors) that capture 
the dominant spectral properties of the original matrix. The singular values and vectors are then extracted from the tridiagonal matrix using 
standard eigenvalue techniques like QR iteration.
\end{textblock}

\begin{textblock}{7.0}(8,1.0)
\Head{Algorithms}\\
Now we can write down pseudocode for both versions of Neon.

{\footnotesize{}
\noindent\rule[-5pt]{.8\textwidth}{0.4pt}\\
{\bf Algorithm 1} Nuclear-Neon update step for linear layer\\
\noindent\rule[5pt]{.8\textwidth}{0.4pt}
\vspace{-15pt}
\noindent\begin{tabbing}
    {\bf Input:} $\lambda$, gradient-like matrix $G$. \\
    {\bf Output:} Weight matrix update $\delta W$.\\*[\smallskipamount]
    \qquad \= 1.\ $U, \Sigma, V := \textrm{Lanczos}(G, 1)$. \\
    {\bf Return} $-\frac{1}{2\lambda}U \Sigma V^T$.
\end{tabbing}
\vspace{-8pt}
\noindent\rule[10pt]{.8\textwidth}{0.4pt}
}

For $F*$-Neon it is a bit trickier~\eqref{eqn:update_F_star}, we need to compute $\tau$ and know number of singular values larger then $\tau$, which we denote by $r$. Assuming that singular values spectrum of $G$ changes little between iterations (which was true in our experiments) we propose the following algorithm.

{\footnotesize{}
\noindent\rule[-5pt]{.8\textwidth}{0.4pt}\\
{\bf Algorithm 2} $F*$-Neon update step for linear layer\\
\noindent\rule[5pt]{.8\textwidth}{0.4pt}
\vspace{-15pt}
\noindent\begin{tabbing}
    \qquad\=\qquad\=\qquad\=\kill % Set three tab stops (4em total)
    {\bf Input:} $\lambda$, $r$, gradient-like matrix $G$. \\
    {\bf Output:} Weight matrix update $\delta W$.\\*[\smallskipamount]
    \> 1. $U, \Sigma, V := \textrm{Lanczos}(G, r + 1)$. \\
    \> 2. $s := \sum_{i=1}^{r}\sigma_r$. \\
    \> 3.\ {\bf If} $(r + 1)\sigma_{r + 1} > s${\bf:}\\
    \>\> 4. $r := r + 1$. 5. $\tau := \frac{s + \sigma_{r}}{r + 1}$.\\
    \> 6.\ {\bf Else if} $(r + 1)\sigma_{r - 1} < s${\bf:}\\
    \>\> 7. $r := r - 1$. 8. $\tau := \frac{s - \sigma_{r}}{r + 1}$.\\
    \> 9.\ {\bf Else:}\\
    \>\> 10. $\tau := \frac{s}{r + 1}$\\
    \> 11. $D = [\Sigma - \tau I]_+$. \\*[\smallskipamount]
    {\bf Save} $r$ for the next iteration.\\
    {\bf Return} $-\frac{1}{\lambda}U D V^T$.
\end{tabbing}
\vspace{10pt}
}
%\noindent\rule[10pt]{.8\textwidth}{0.4pt}

\medskip
\Head{MLP tests}\\
We test Muon, SGD, and Neon on a simple MLP (2 linear layers) that solves CIFAR-10 classification problem. Muon converges faster than Neon and reaches higher accuracy.
\begin{figure}[h]
    \center{\includegraphics[width=0.7\linewidth]{../figs/mlp0605/loss_vs_epochs_mlp.pdf}}
    \caption{MLP validation loss}
    \label{fig:mlp_loss}
\end{figure}

\begin{figure}[h]
    \center{\includegraphics[width=0.7\linewidth]{../figs/mlp0605/accuracy_vs_time_mlp.pdf}}
    \caption{MLP Accuracy vs wallclock time}
    \label{fig:mlp_loss}
\end{figure}

\end{textblock}

\begin{textblock}{7.0}(16,0.75)
\Head{CIFAR-10 tests}\\
Now we compare Muon and Neon on ResNet from Keller's cifar10-aribench at GitHub. On RTX-4050, Muon is approximately twice as fast as Neon. That is likely caused by ineffective implementation of Neon's update, because in theory updates of Neon are more light and simple, especially for Nuclear Neon. Interestingly, Neon has a ceiling of convergence about 70\% which is not observed for Muon. Moreover, Neon requires far smaller learning rates to converge. All these aspects should be considered when we improve the algorithm.

    \begin{figure}[h]
        \center{\includegraphics[width=0.75\linewidth]{../figures/13may_muon_neon/test_acc_vs_epoch.pdf}}
        \caption{Accuracy vs Epoch for Muon vs Neon}
        \label{fig:mlp_loss}
    \end{figure}

\medskip
\Head{NanoGPT tests}\\
We finetuned NanoGPT by Alex Karpathy on two NVIDIA RTX 4090 GPUs (24 GB each) on tiny stories dataset via 200 iterations. Both Neon and Muon showed better convergence then AdamW. Neon showed similiar convergence rate to Muon

% \begin{figure}[h]
%     \center{\includegraphics[width=0.8\linewidth]{../../figs/mlp24/loss_val.png}}
%     \caption{NanoGPT validation loss}
%     \label{fig:nanogpt_loss}
% \end{figure} 
% \begin{figure}[h]
%     \center{\includegraphics[width=0.8\linewidth]{../../figs/nanogpt/train_loss.pdf}}
%     \caption{NanoGPT train loss}
%     \label{fig:nanogpt_train_loss}
% \end{figure}
\begin{figure}[h]
    \center{\includegraphics[width=0.8\linewidth]{../figs/nanogpt/val_loss.pdf}}
    \caption{NanoGPT validation loss}
    \label{fig:nanogpt_val_loss}
\end{figure}
\medskip
\Head{Conclusion}\\
A new norm-motivated algorithm proved to be not so effective as Muon, at least from the first sight. However, its updates seem to be much faster to compute, and more so for large matrices that have not been tested yet. We believe our algorithm is not an alternative to Muon, but probably a more effective tool for solving some other optimization problems. A deep inspection of nuclear norm and its properties might give us an insight into Neon's future.

\medskip
%\Head{Acknowledgements}\\
%We thank Jeremy Bernstein for his idea to use norms to derive a optimizer.

\end{textblock}

\begin{textblock}{6}(-0,11.3)
    \footnotesize
    \raggedright
    \textcolor{gray}{\emph{Jeremy Bernstein and Laker Newhouse. Old optimizer, new norm: An anthology (2024)}}
    \end{textblock}

    \begin{textblock}{6}(-0,11.5)
    \footnotesize
    \raggedright
    \textcolor{gray}{\emph{Jeremy Bernstein. Deriving Muon (2025)}}
    \end{textblock}

    \begin{textblock}{6}(-0,11.7)
    \footnotesize
    \raggedright
    \textcolor{gray}{\emph{Keller Jordan et al. Muon: An optimizer for hidden layers in neural networks (2024)}}
    \end{textblock}

% QR code in the lower right corner
\begin{textblock}{2}(21.2,9.6)
\begin{center}
\includegraphics[width=0.7\TPHorizModule]{neon_repo_qr.png}
\end{center}
\end{textblock}

\end{document}